\paperTitle

### 2014-02-05 doc/icml2014/abstract.tex
push

Recent work in method of moments provide consistent estimates
  for latent-variable models, avoiding local optima issues,
  but these methods can only be applied to certain types of graphical models.
In this work, we show that the method of moments in conjunction
  with a composite marginal likelihood objective
  yields consistent parameter estimates for a much broader
  class of directed and undirected graphical models,
  including loopy graphs with high treewidth.
%In particular, our algorithm applies to any graphical model wherein each
  %hidden variable has at least three conditionally independent observed
  %variables.
  %that sidestep local optima issues for a general class of
  %directed and undirected graphical models.
%Parameter estimation in latent-variable models based on EM
  %is prone to local optima.
  %due to the non-convexity of the marginal log-likelihood. 
%In this paper, we extend recent advances in the method of moments 
  %to provide consistent parameter estimates
  %that sidestep local optima issues for a general class of
  %directed and undirected graphical models.
%In particular, our algorithm applies to any graphical model wherein each
  %hidden variable has at least three conditionally independent observed
  %variables.
%This family includes loopy graphs with high treewidth.
Specifically, we use tensor factorization to reveal partial information
about the hidden variables, rendering the otherwise non-convex negative
log-likelihood convex.
%We use the method of moments to tie each hidden to some observed
  %variables, and show that the rest of the problem can be cast as convex
  %optimization of the likelihood.
Our approach gracefully extends to models outside our class by
  incorporating the partial information via posterior regulraization.
\end{abstract}

\paragraph{Last updated:} \today, \currenttime

\paragraph{Errata / Incomplete}
\begin{enumerate}
  \item Use \texttt{\textbackslash left(}.
  \item Use $\ell$ instead of $L$.
  \item convex and negative log likelihood.
\end{enumerate}

>>>>

Recent work on the method of moments provide consistent estimates
  for latent-variable models, avoiding local optima issues,
  but these methods are applicable only to certain types of graphical models.
  On the other hand, pure likelihood objectives are difficult to optimize.
In this work, we show that using the method of moments in conjunction
  with composite marginal likelihood
  yields consistent parameter estimates for a much broader
  class of discrete directed and undirected graphical models,
  including loopy graphs with high treewidth.
%In particular, our algorithm applies to any graphical model wherein each
  %hidden variable has at least three conditionally independent observed
  %variables.
  %that sidestep local optima issues for a general class of
  %directed and undirected graphical models.
%Parameter estimation in latent-variable models based on EM
  %is prone to local optima.
  %due to the non-convexity of the marginal log-likelihood. 
%In this paper, we extend recent advances in the method of moments 
  %to provide consistent parameter estimates
  %that sidestep local optima issues for a general class of
  %directed and undirected graphical models.
%In particular, our algorithm applies to any graphical model wherein each
  %hidden variable has at least three conditionally independent observed
  %variables.
%This family includes loopy graphs with high treewidth.
Specifically, we use tensor factorization to reveal partial information
about the hidden variables.  This renders the otherwise non-convex
likelihood objective convex, and EM can then be used to attain the global optimum.
%We use the method of moments to tie each hidden to some observed
  %variables, and show that the rest of the problem can be cast as convex
  %optimization of the likelihood.
%Our approach gracefully extends to models outside our class by
  %incorporating the partial information via posterior regulraization.
\end{abstract}

%\paragraph{Last updated:} \today, \currenttime
%
%\paragraph{Errata / Incomplete}
%\begin{enumerate}
%  \item Use \texttt{\textbackslash left(}.
%  \item Use $\ell$ instead of $L$.
%  \item convex and negative log likelihood.
%\end{enumerate}
%

### 2014-02-05 doc/icml2014/discussion.tex

% Correlation structure
Other authors, e.g \citet{anandkumar12lda, anandkumar2013linear,
  halpern13noisyor}, learn parameters for models outside this
  family, they do so with some knowledge of the correlation structure
  between hidden variables. 
In a pre-processing step, they ``subtract'' out the correlations and
  recover a graph that satisfies the bottleneck property
  (\propertyref{bottleneck}).
% TOO BOLD
%Thus, while \propertyref{bottleneck} is much stronger than
%  \propertyref{exclusive-views}, we expect that without any knowledge of
%  the correlation between variables, \propertyref{bottleneck} is
%  necessary for identifiability.
%Our work uses the \TensorFactorize algorithm presented by
%  \citet{anandkumar13tensor}, first
%  studied the consistent parameter estimation algorithm for the
%  three-view mixture model.
%
%\todo{What about Mossel and Roch?}
For example,
\citet{anandkumar12lda} are able to recover parameters for correlated
  discrete mixtures, e.g. latent Dirichlet allocation; their approach
  follows that of the three-view mixture model, with a pre-processing step
  that subtracts the correlation between the views.
\citet{anandkumar2013linear} study a special family of Bayesian networks
  in which the observed variables are linearly related to a sparse set of
  hidden variables. Once again, the key step is to identify the
  correlations between observed variables and exclude them from the views.
\citet{halpern13noisyor} propose an algorithm to recover factors for
  a more densely connected bipartite noisy-or network; they also use
  bottlenecks to learn parameters, but exploit properties of the model
  to negate correlations from learned edges, allowing for more
  bottlenecks to be discovered. 
Our work makes a different tradeoff: we do not assume any special correlation structure,
but do rely on every hidden variable having an exclusive view so that we can
learn the underlying structure.

@@

% Observable operator
Another prominent line of work in the method of moments community has
  focused on recovering an {\em observable operator
  representation} of the observations.
  These methods which allows for prediction of new observations, but does not
  identify the actual parameters of the model, which makes them difficult to use
  in conjunction with likelihood-based models which are driven by parameters.
\citet{song2011spectral} propose an algorithm to learn an observable
  operator representations for latent tree graphical models, like the
  one in \figureref{examples-tree}, assuming the bottleneck property (\propertyref{bottleneck}). 
Their approach is similar to our first step of learning conditional moments,
  but only consider trees.
\citet{parikh12spectral} extend this approach to general graphical
  models in which \propertyref{bottleneck} holds, but consider the latent
  junction tree representation. 
  Consequently, the size of the observable representations is exponential in
  the treewidth.
  In contrast, our algorithm only constructs moments of the order of size of the cliques
  (and neighborhoods for pseudolikelihood), which can be much smaller.
For example, an $n\times n$ grid model has a treewidth of $n$, but each
  clique is of size at most $2$ and the degree is at most $2$.
%In some sense, an observable operator re-formulation of our algorithm
  %subsumes this work.

>>>>

% Correlation structure
Other authors, e.g \citet{anandkumar12lda, anandkumar2013linear,
  halpern13noisyor}, learn parameters for models outside this
  but family, but they do so with some knowledge of the correlation structure
  between hidden variables. 
In a pre-processing step, they ``subtract'' out the correlations and
  recover a graph that satisfies the bottleneck property
  (\propertyref{bottleneck}).
% TOO BOLD
%Thus, while \propertyref{bottleneck} is much stronger than
%  \propertyref{exclusive-views}, we expect that without any knowledge of
%  the correlation between variables, \propertyref{bottleneck} is
%  necessary for identifiability.
%Our work uses the \TensorFactorize algorithm presented by
%  \citet{anandkumar13tensor}, first
%  studied the consistent parameter estimation algorithm for the
%  three-view mixture model.
%
%\todo{What about Mossel and Roch?}
For example,
\citet{anandkumar12lda} are able to recover parameters for correlated
  discrete mixtures, e.g. latent Dirichlet allocation; their approach
  follows that of the three-view mixture model, with a pre-processing step
  that subtracts the correlation between the views.
\citet{anandkumar2013linear} study a special family of Bayesian networks
  in which the observed variables are linearly related to a sparse set of
  hidden variables. Once again, the key step is to identify the
  correlations between observed variables and exclude them from the views.
\citet{halpern13noisyor} propose an algorithm to recover factors for
  a more densely connected bipartite noisy-or network; they also use
  bottlenecks to learn parameters, but exploit properties of the model
  to negate correlations from learned edges, allowing for more
  bottlenecks to be discovered. 
Our work makes a different tradeoff: we do not assume any special correlation structure,
but do rely on every hidden variable having an exclusive view so that we can
learn the underlying structure.

@@

% Observable operator
Another prominent line of work in the method of moments community has
  focused on recovering an {\em observable operator
  representation} of the observations.
  These methods which allows for prediction of new observations, but do not
  identify the actual parameters of the model, which makes them difficult to use
  in conjunction with likelihood-based models. % which are driven by parameters.
\citet{song2011spectral} propose an algorithm to learn an observable
  operator representations for latent tree graphical models, like the
  one in \figureref{examples-tree}, assuming the bottleneck property (\propertyref{bottleneck}). 
Their approach is similar to our first step of learning conditional moments,
  but only consider trees.
\citet{parikh12spectral} extend this approach to general graphical
  models in which \propertyref{bottleneck} holds, but consider the latent
  junction tree representation. 
  Consequently, the size of the observable representations is exponential in
  the treewidth.
  In contrast, our algorithm only constructs moments of the order of size of the cliques
  (and neighborhoods for pseudolikelihood), which can be much smaller.
For example, an $n\times n$ grid model has a treewidth of $n$, but each
  clique is of size at most $2$ and the degree is at most $4$.
%In some sense, an observable operator re-formulation of our algorithm
  %subsumes this work.

### 2014-02-05 doc/icml2014/undirected.tex

% FIGURE: example
\begin{figure}
  \label{fig:examples-mrf}
  \centering
  \includegraphics[width=0.6\columnwidth]{figures/mrf.pdf}
%  \subimport{figures/}{mrf.tikz}
  \caption{Example: undirected grid model where each hidden variable has two
  conditionally independent observations.
  We derive a consistent estimate.}
\end{figure}

@@

%In this section, we will show that we can extend the results to
  %undirected graphical models parameterized as log-linear latent variable
  %models.
On the other hand, 
if we were able to observe $\vh$, the \emph{supervised} likelihood would be concave:
\begin{align}
\label{eqn:logLinearSupervised}
L_\text{sup}(\theta) &\eqdef \E_{(\vx,\vh) \sim \sD_\text{sup}}[\log p_\theta(\vx,\vh)] \\
                     &= \theta^\top \left(\sum_{\sC \in \sG} \E[\phi(\vx_\sC,\vh_\sC)]\right) - A(\theta).
\end{align}
While we don't have supervised data,
note that the method of moments gives us latent marginals $Z_\sC = \BP(\vh_\sC)$.
Furthermore, given $\mOpp{v}{i}$ and $Z_\sC$ of any $\sC \ni v$, we can recover
all marginals $Z_\sC = \BP(\vx_\sC, \vh_\sC)$ (with a slight augmentation of notation).
Finally, the key step is that we can compute all the expected feature vectors from the marginals:
\begin{align}
\label{eqn:logLinearFeatures}
\mu_\sC \eqdef \E[\phi(\vx_\sC,\vh_\sC)] = \sum_{\vx_\sC,\vh_\sC} Z_\sC(\vx_\sC,\vh_\sC) \phi(\vx_\sC,\vh_\sC).
\end{align}
Therefore, we can optimize the supervised likelihood objective without actually
having supervised data.
Of course, in the finite data regime, the method of moments yields estimates
$\hat \mu^\text{mom}_\sC$ which approximate the true $\mu_\sC$.
In supervised learning, we have different estimate $\hat\mu^\text{sup}_\sC$ of $\mu_\sC$ based on an empirical average
over the data points.
In the limit of infinite data, both estimators converge to $\mu_\sC$.

@@

\paragraph{Pseudolikelihood.}
Unlike for directed models, in the undirected case,
graphs with high treewidth present a challenge
for even fully-supervised maximum likelihood due to the intractability of the log-partition function.
This motivates looking at pseudolikelihood objective of \citep{besag75pseudo} for a computationally more
efficient solution.
The pseudolikelihood objective sums over the log-probability of each variable $a$ given its neighbors $\sN(a)$:
$$L_\text{pseudo}(\theta) \eqdef \E_{x \sim \sD}\left[\sum_{a \in X \cup H} \log p_\theta(a \mid x_{\sN(a)})\right].$$
In the fully-supervised setting, pseudolikelihood provides computationally efficient consistent estimates,
albeit at the cost of statistical efficiency.
Coincidentally, the same high-level motivation for using method of moments in the first place.

@@

In general, we might only have exclusive views (\propertyref{exclusive-views}) for a subset of the cliques.
In this case, we can estimate the expected features $\mu_\sC$ for those cliques
and use posterior regularization \citep{graca08em} or generalized expectation criteria \citep{mann08ge}
to encourage $\E_{p_\theta}[\phi(\vx_\sC,\vh_\sC)]$ to match $\mu_\sC$.
The resulting objective functions are still non-convex, but in practice we observed
that the extra information $\mu_\sC$ helps alleviate local optima.
%We can then integrate these constraints using posterior regularization \citep{graca08em}:
%\begin{align}
%L_\text{pr}(\theta) \eqdef L_\text{unsup}(\theta) + C\min_{q \in \sQ} \text{KL}(q(\vx,\vh) \| p_\theta(\vx,\vh)).
%\end{align}
%The main idea is to regularize the parameters towards values that produce the 
%While the optimization problem \eqref{eqn:obj-ml} is non-convex, we
  %expect that the measurements aid recover of parameters, and indeed see
  %this is true for experiments on simulated data.
%\todo{Include figure}

>>>>

% FIGURE: example
\begin{figure}
  \centering
  \includegraphics[width=0.6\columnwidth]{figures/mrf.pdf}
%  \subimport{figures/}{mrf.tikz}
  \caption{Example: undirected grid model where each hidden variable has two
  conditionally independent observations.
  We derive a consistent estimate.}
  \label{fig:examples-mrf}
\end{figure}

@@

%In this section, we will show that we can extend the results to
  %undirected graphical models parameterized as log-linear latent variable
  %models.
On the other hand, 
if we were able to observe $\vh$, the \emph{supervised} likelihood would be concave:
\begin{align}
\label{eqn:logLinearSupervised}
L_\text{sup}(\theta) &\eqdef \E_{(\vx,\vh) \sim \sD_\text{sup}}[\log p_\theta(\vx,\vh)] \\
                     &= \theta^\top \left(\sum_{\sC \in \sG} \E[\phi(\vx_\sC,\vh_\sC)]\right) - A(\theta).
\end{align}
While we don't have supervised data,
note that the method of moments gives us hidden marginals $Z_\sC = \BP(\vh_\sC)$.
Furthermore, given $\mOpp{v}{i}$ and $Z_\sC$ of any $\sC \ni v$, we can recover
all marginals $Z_\sC = \BP(\vx_\sC, \vh_\sC)$ (with a slight augmentation of notation).
Finally, the key step is that we can compute all the expected feature vectors from the marginals:
\begin{align}
\label{eqn:logLinearFeatures}
\mu_\sC \eqdef \E[\phi(\vx_\sC,\vh_\sC)] = \sum_{\vx_\sC,\vh_\sC} Z_\sC(\vx_\sC,\vh_\sC) \phi(\vx_\sC,\vh_\sC).
\end{align}
Therefore, we can optimize the supervised likelihood objective without actually
having supervised data.
Of course, in the finite data regime, the method of moments yields estimates
$\hat \mu^\text{mom}_\sC$ which approximate the true $\mu_\sC$.
In supervised learning, we have different estimate $\hat\mu^\text{sup}_\sC$ of $\mu_\sC$ based on an empirical average
over the data points.
In the limit of infinite data, both estimators converge to $\mu_\sC$.

@@

\paragraph{Pseudolikelihood.}
Unlike for directed models, in the undirected case,
graphs with high treewidth present a challenge
for even fully-supervised maximum likelihood due to the intractability of the log-partition function.
This motivates looking at pseudolikelihood objective of \citep{besag75pseudo} for a computationally more
efficient solution.
The pseudolikelihood objective sums over the log-probability of each variable $a$ given its neighbors $\sN(a)$:
$$L_\text{pseudo}(\theta) \eqdef \E_{x \sim \sD}\left[\sum_{a \in X \cup H} \log p_\theta(a \mid x_{\sN(a)})\right].$$
In the fully-supervised setting, pseudolikelihood provides computationally efficient consistent estimates,
albeit at the cost of statistical efficiency.
Coincidentally, this is the same high-level motivation for using method of moments in the first place.

@@

In general, we might only have exclusive views (\propertyref{exclusive-views}) for a subset of the cliques.
In this case, we can estimate the expected features $\mu_\sC$ for those cliques
  and use posterior regularization \citep{graca08em} or generalized expectation criteria \citep{mann08ge}
  to encourage $\E_{p_\theta}[\phi(\vx_\sC,\vh_\sC)]$ to match $\mu_\sC$.
The resulting objective functions are still non-convex, but we expect
  local optima to be alleviated.

%but in practice we observed that the extra information $\mu_\sC$ helps alleviate local optima.
%We can then integrate these constraints using posterior regularization \citep{graca08em}:
%\begin{align}
%L_\text{pr}(\theta) \eqdef L_\text{unsup}(\theta) + C\min_{q \in \sQ} \text{KL}(q(\vx,\vh) \| p_\theta(\vx,\vh)).
%\end{align}
%The main idea is to regularize the parameters towards values that produce the 
%While the optimization problem \eqref{eqn:obj-ml} is non-convex, we
  %expect that the measurements aid recover of parameters, and indeed see
  %this is true for experiments on simulated data.
%\todo{Include figure}
